Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-01-20 docs/ACM/paper.tex
Adding some text in the intro of Section 3.

%% Packages that we are using in this version of the work:
\usepackage{hyperref}
\usepackage{amsmath}

@@

\paragraph{The syntax that we handle.}
Our goal is to identify {\em tasks} in {\em structured} programs.
Definition~\ref{def:task} formalizes the notion of task.
A structured program is a program that can be partitioned in {\em hammock
regions}, a concept introduced by Ferrante {\em et al.}~\cite{Ferrante87} in
the middle 80's.
The structure of this program is described by a a Control Flow Graph $G$ --
a directed graph with a starting node $s$, and an exit node $x$.
A hammock region is a subgraph $G' \subseteq G$, with a special node
$h$, which {\em dominates}\footnote{A node $n_1 \in G$ dominates another node
$n_2 \in G$ if every path from  $S$ to $n_2$ goes across $n_1$.
Inversely, $n_1$ post-dominates $n_2$ if every path from $n_2$ to $x$ must
go across $n_1$.} every other node in $G'$.
Additionally, there exists a node $w \in G$, $w \notin G'$, such that
$w$ {\em post-dominates} every node in $G'$.
In this definition, $h$ is the {\em entry point}, and $w$ is the {\em exit point}
of $G'$.

@@

Definition~\ref{def:task} uses the concept of {\em memory region}.
Syntactically, memory regions are described by program variables and/or pointers
plus ranges of offsets that these pointers are allowed to dereference.
Given two tasks: $T_1 = (G_1, M_{i1}, M_{o1})$ and $T_2 = (G_2, M_{i2}, M_{o2})$,
if $M_{o1} \cap M_{i2} \neq \emptyset$, then we say that $T_2$ {\em depends} on
$T_1$.
If a program $P$ is partitioned into a set of $n$ tasks, then this partition is
said to be {\em correct} if: (i) it does not contain cyclic dependence relations;
and (ii) any topological ordering of the tasks, as determined by dependence
relations, leads to the same results as the sequential execution of $P$.

\paragraph{The annotations that we use.}

>>>>

%% Packages that we are using in this version of the work:
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{paralist}

@@

\paragraph{Tasks: syntax and scope:}
Our goal is to identify {\em tasks} in {\em structured} programs.
Definition~\ref{def:task} formalizes the notion of task.
A structured program is a program that can be partitioned in {\em hammock
regions}, a concept introduced by Ferrante {\em et al.}~\cite{Ferrante87} in
the middle 80's.
The structure of this program is described by a a Control Flow Graph $G$ --
a directed graph with a starting node $s$, and an exit node $x$.
A hammock region is a subgraph $G' \subseteq G$, with a special node
$h$, which {\em dominates}\footnote{A node $n_1 \in G$ dominates another node
$n_2 \in G$ if every path from  $S$ to $n_2$ goes across $n_1$.
Inversely, $n_1$ post-dominates $n_2$ if every path from $n_2$ to $x$ must
go across $n_1$.} every other node in $G'$.
Additionally, there exists a node $w \in G$, $w \notin G'$, such that
$w$ {\em post-dominates} every node in $G'$.
In this definition, $h$ is the {\em entry point}, and $w$ is the {\em exit point}
of $G'$.

@@

Definition~\ref{def:task} uses the concept of {\em memory region}.
Syntactically, memory regions are described by program variables and/or pointers
plus ranges of offsets that these pointers are allowed to dereference.
Given two tasks: $T_1 = (G_1, M_{i1}, M_{o1})$ and $T_2 = (G_2, M_{i2}, M_{o2})$,
if $M_{o1} \cap M_{i2} \neq \emptyset$, then we say that $T_2$ {\em depends} on
$T_1$.
If a program $P$ is partitioned into a set of $n$ tasks, then this partition is
said to be {\em correct} if it abides by Definition~\ref{def:correctness}.

\begin{definition}[Correctness]
\label{def:correctness}
A set $T$ of $n$ tasks is a correct parallelization of a program $P$ if:
\begin{compactitem}
\item $T$ does not contain cyclic dependence relations;
\item the execution of the tasks in $T$ in any ordering determined by dependence
relations, leads to the same results as the sequential execution of $P$.
\end{compactitem}
\end{definition}

\begin{example}[Memory Regions]
\label{ex:regions}
\fer{XX Add an example to describe the notions of regions and dependences.
The best is to modify one of the programs in Section 2, so that we can refer to
that program.
Pedro, can you think about it and propose something?}
\end{example}

\paragraph{The annotation zoo.}
Our equipment of choice to reveal task parallelism in programs is OpenMP 3.0.
This annotation system provides us with a rich set of pragmas that we can use
to indicate to compiler and runtime environment which tasks must be created.
We describe below which annotations we are using, and explain, informally, their
semantics.
A full overview of the syntax and semantics of these annotations is publicly
available\footnote{For a quick overview, we refer the reader to the leaflet
``Summary of OpenMP 3.0 C/C++ Syntax", which is made available by the OpenMP
group.}.

\begin{compactitem}
\item \textsf{task} {\em clauses}: creates a new task, according to the
parameters specified in the {\em clauses} list:
\begin{compactitem}
\item \textsf{default([shared/private])}: indicates that variables are either
shared or replicated among tasks.
\item \textsf{firstprivate($v$)}: indicates that $v$ must be replicated among
tasks, and initialized with its value at the point where the annotation is
executed.
\item \fer{Pedro/Gleison: can you guys continue this list?}
\end{compactitem}
\item \textsf{taskwait}: forces tasks to synchronize at the point of annotation.
\end{compactitem}
